\section{Hyper Regular Expressions and Application in Security}

It is straightforward to see that the underlying NFM of an NHF defines a 
regular expression over hyperwords. Such a regular expression along with the 
quantification of the NHF constitutes a {\em hyper regular expression} 
(HRE). For example, consider the left NHF $\A_1$. The HRE of this NHF is the 
following:
%
$$
\forall x_1\forall x_2\Big((a, a) \mid (b, b)\Big)^*\Big((\#, b)^* \mid (b, 
\#)^* \Big)
$$
%
We now show the application of HREs in specifying information-flow 
security policies.

{\em Noninteference}~\cite{gm82} requires that commands issued by users holding 
high clearances be removable without affecting observations of users holding 
low clearances:
%
$$
\varphi_{\mathsf{ni}} = \forall x_1.\exists x_2.(l, l\lambda)^*
$$
%
where $l$ denotes a low state and $l\lambda$ denotes a low state where all high 
commands are replaced by a dummy value $\lambda$.

{\em Observational determinism}~\cite{zm03} requires that if two executions of 
a system start in low-equivalent states, then these execution should remain low 
equivalent:
%
$$
\varphi_{\mathsf{od}} = \forall x_1.\forall x_2. (l, l)^+ + (\bar{l}, 
\bar{l})(\$, \$)^* + (l, \bar{l})(\$, \$)^* + (\bar{l}, l)(\$, \$)^*
$$
%
where $l$ denotes a low state, $\bar{l} \in \Sigma - \{l\}$, and $\$ \in 
\Sigma$. We note that similar polices such as {\em Boudol and Castellani’s 
noninterference}~\cite{bd02} can be formulated in the same 
fashion\footnote{This policy states that any two executions that start from 
bisimilar states (in terms of memory low-observability, should remain 
bisimilarly low-observable.}.

{\em Generalized noninterference} (GNI)~\cite{m88} allows nondeterminism in 
the low-observable behavior, but requires that low-security outputs may 
not be altered by the injection of high-security inputs:
%
$$
\varphi_{\mathsf{gni}} = \forall x_1.\forall x_2.\exists x_3. \bigg((h, l, hl) 
+ (\bar{h}, l, \bar{h}l) + (h, \bar{l}, h\bar{l}) + (\bar{h}, \bar{l}, 
\bar{h}\bar{l}) \bigg)^*
$$
%
where $h$ denotes the high input, $l$ denotes the low output, $\bar{l} \in 
\Sigma -\{l\}$, and $\bar{h} \in \Sigma - \{h\}$.

{\em Declassification}~\cite{ss00} relaxes noninterference by allowing leaking 
information when necessary. Some programs need to reveal secret information to 
fulfill functional requirements. For example, a password checker must reveal 
whether the entered password is correct or not:
%
$$
\varphi_{\mathsf{dc}} = \forall x_1.\forall x_2. (li,li)(pw, pw)(lo, lo)^+
$$
%
where $li$ denotes low-input state, $pw$ denotes that the password is correct, 
and $lo$ denotes low-output states. We note that for brevity, in the above 
formula, we do not include behaviors where the first two events are not low or 
in the second event, the password in not valid. 

{\em Termination-sensitive noninterference} requires that for two executions 
that
 start from low-observable states, information leaks are not permitted by 
the program's termination behavior:
%
$$
\varphi_{\mathsf{tsni}} = \forall x_1.\forall x_2. (l, l)(\$, \$)^*(l, l)  + 
(\bar{l}, \bar{l})(\$, \$)^* + (l, \bar{l})(\$, \$)^* + (\bar{l}, l)(\$, \$)^*
$$
where $l$ denotes a low state and $\$ \in \Sigma$.

